---
title: "project"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-05-05"
---
```{r}
library(tidyverse)
library(conflicted)  

library(tidyverse)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```

```{r}

project <- read.table("C:\\Users\\rragh\\Downloads\\project.csv", header=TRUE, sep=",")

```

> Requirement 1:

```{r}
# We used the data frame to review the data
data.frame(project)

print(project)
```
```{r}
project <- as_tibble(project)
project

```


```{r}
# Dimension of the data 
dim(project)
```

```{r}
# To summarize the data frame into just one value or vector
summary(project)
```
 
```{r}
# Return a list of objects and their structure.
str(project)
```

> This data is about education employees, and it contains four values, which are Id, Rank, Experience and Salary. Out of str function the types of our dataset are: Numeric for ID, Character for Rank, Numeric for experience, and Numeric for salary, All data have the same range which is [ 1 : 403 ] We also have 400 employees with varying experience, and the salary starts from $57,800 Also the experiences starts from 1 year.



> Requirement 2:

```{r}
# for null values
colSums(is.na(project))
```


```{r}
#Sum the null values
sum(colSums(is.na(project)))
```


```{r}
# for outlier  
summary(project$Experience)
```

```{r}
summary(project$Salary)
```


```{r}
#Show the outlier in boxplot
boxplot(project$Experience,
  ylab = "Experience"
)
```

```{r}
boxplot.stats(project$Experience)$out
```
> The range of the experience is 60 and the outlier is 150

```{r}
#Show the outlier in boxplot
boxplot(project$Salary,
  ylab = "salary"
)
```

```{r}
boxplot.stats(project$Salary)$out
```
> The range of the salary is 204000-205500 and the outlier is 231545 

```{r}
#Sum the duplicated values
sum(duplicated(project$ID))
```
> This is the 3 duplicated values

```{r}
# We check the Rank column if there are errors in writing the names or not
name <- project %>%
  filter(Rank != "AssocProf") %>%
  filter(Rank != "AsstProf") %>%
  filter(Rank != "Prof") %>%
  group_by(Rank)

```


> Requirement 3:


```{r}
#remove duplicated
project <- project[!duplicated(project$ID),]
```


```{r}

sum(duplicated(project$ID))
```



```{r}
#remove missing values
project <- na.omit(project)
```

```{r}
colSums(is.na(project))
```

```{r}
# remove outliers salary
quartiles <- quantile(project$Salary, probs=c(.25, .75))
IQR <- IQR(project$Salary)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
 
data <- subset(project, project$Salary > Lower & project$Salary < Upper)
 
# Dimension of the data after removing the outlier 
dim(data)
```

```{r}
#Show the salary boxplot after remove the outlier
boxplot(data$Salary,
  ylab = "salary"
)
```

```{r}
# remove outliers Experience
quartiles <- quantile(data$Experience, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(data$Experience)
 
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
 
data2 <- subset(data, data$Experience > Lower & data$Experience < Upper)
 
# Dimension of the data after removing the outlier
dim(data2)
```

```{r}
#Show the Experience boxplot after remove the outlier
boxplot(data2$Experience,
  ylab = "Experience"
)
```

```{r}
# changing wrong names
data2 <- mutate(data2, Rank = recode(.x=Rank, "AssocProff"="AssocProf"))
```

```{r}
# changing wrong names
data2 <- mutate(data2, Rank = recode(.x=Rank, "AssstProf"="AsstProf"))
```


```{r}
# show data after preprocessinge
data2
```

```{r}
# show summary after preprocessinge
summary(data2)

```



> Requirement 4:

```{r}
library(ggplot2)
```

```{r}
# Bar chart for Faculty_Rank
ggplot(data2, aes(x=Rank)) + geom_bar() + labs(title="Faculty Rank Distribution")

```

```{r}
# Histogram for Salary
ggplot(data2, aes(x=Salary)) + geom_histogram(binwidth=5000) + labs(title="Faculty Salary Distribution")

```


```{r}
# Histogram for Experience
ggplot(data2, aes(x=Experience)) + geom_histogram(binwidth=1) + labs(title="Faculty Experience Distribution")
```


> Requirement 5:

```{r}
# Scatterplot for Salary vs Experience colored by Faculty_Rank
ggplot(data2, aes(x=Experience, y=Salary, color=Rank)) + geom_point() + geom_smooth(method = "lm") +labs(title="Faculty Salary vs Experience by Rank")
```


```{r}
ggplot(data = data2, aes(x = Rank, y = Salary,color=Rank)) +
  geom_violin()
```


```{r}
ggplot(data2, aes(x=Rank, y=Experience,color=Rank)) +
 
  geom_violin()
```


```{r}
ggplot(data2, aes(y=Salary, x=Experience)) +
 
  geom_violin()
```


```{r}
base.plot <- ggplot(data2, aes(x = Rank)) +
 xlab("Rank")
base.plot + geom_density()
```


```{r}
base.plot + geom_density(aes(fill = Rank), alpha = 0.6)
```

> Requirement 6:

```{r}
#"rank" affect "salary"
mean_Salary <- data2 %>% group_by(Rank) %>% 
  summarise(mean(Salary))

```


```{r}
# Experience impact salary
max_Experience_Salary  <- data2 %>% group_by(Rank) %>% 
  summarise(max(Salary), max(Experience))
```

```{r}
# gaps for employees with the same experience and rank
gaps  <- data2 %>% group_by(Experience) %>% 
  summarise(max(Salary),max(Rank))
```

```{r}
head(gaps)
```

